---
title: "Homework: Data Visualization"
author: "Firstname Lastname"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 3)
```

```{r load-sas-libraries, echo = F, message = F, warning = F}

library(SASmarkdown)
sas_enginesetup(sashtml=sashtml)

sasexe <- "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"
sasopts <- "-nosplash -ls 75"

# Linux options (for grading, please leave this in!)
# sasexe <- "/usr/local/SASHome/SASFoundation/9.4/bin/sas_en"
# sasopts <- "-ls 75"
```

## Instructions

When you are finished with the assignment:

1. Save the file as 08_Lastname_Firstname.Rmd and compile it
2. Commit the Rmd file and corresponding html file to your homework git repo
3. Push the commit to github
4. Locate the Rmd file on github and click on the SHA link for the file
5. Paste the url of the resulting page into Canvas to submit your assignment.

Your assignment must compile on a different computer as long as the saspath is set correctly for that machine.
This means you will need to use a local file path when you read the data in via R (and SAS).

## Swiss Banknotes

The R package `alr3` contains a data set called banknote, consisting of (physical) measurements on 200 Swiss bank notes, 100 of which are genuine, while the other half is counterfeit. Load this data set (you might have to install the package) using the code below. 

```{r, echo = F}
# install.packages("alr3")
library(alr3) # if this throws an error of the form 'there is no package called alr3', uncomment the line above, run it once, then comment the line out again and run the code chunk again.
data(banknote)
banknote$Y <- factor(banknote$Y, labels = c("Genuine", "Fake"))

library(ggplot2)
```

For each of the three figures you create below, write a two-three sentence summary, describing the structure of the plot: what type of plot is it? Which variables are mapped to x, to y, and to the (fill) color? What is the main message of the plot: what is your main finding, i.e. what do you want viewers to learn from the plot? Are there any anomalies or outliers?

1. Use one of the object inspecting functions and describe the data set - what do the variables appear to mean?
```{r}
```


2. Use `ggplot2` to draw a barchart of Y (0 is genuine, 1 is counterfeit). Also, map Y to the fill color of the bars. (Yes, this plot is a bit simplistic, but what does it show?)
```{r}

```

3. Use `ggplot2` to draw a histogram of one of the variables in the dataset that shows a distinction between genuine and counterfeit banknotes. Use fill color to show this difference. Choose the binwidth such that there are no gaps in the middle range of the histogram.
```{r}
```

4. Use `ggplot2` to draw a scatterplot of two (continuous) measurements, color by Y. Try to find a pair of measurements that provides as much separation as possible between the clusters of points for genuine and counterfeit banknotes.
```{r}
```

## Take everything you know, and use it for evil

The textbook spent lots of time showing you how to create different types of graphics, but I spent a lot less time showing you all of the different ways you could customize graphics in both ggplot2 and SAS. In this problem, I want you to create the *ugliest* version of the following graphs you can, and then **explain why**, exactly, you made the decisions you did, and which principles of good graphics you've intentionally violated. 

Ugliness is subjective, so the goal here is for you to explore the different ways you can customize the finer details of ggplot2 and SAS graphics. Make sure your finished masterpiece has appropriate axis labels and a title (after all, even ugly plots need to be correctly labeled!). You are free to add additional variables and layers, modify the aesthetics used, and leverage other packages. I have provided this code as a starting point, not as a way to limit your creativity. [If you need inspiration, look here.](https://www.allisonhorst.com/post/do-your-worst/)

### Part 1: R
I have provided some basic code to get you started. In R, we'll use the `palmerpenguins` package, which has [fantastic documentation with really awesome artwork](https://allisonhorst.github.io/palmerpenguins/). 

```{r}
if (!"palmerpenguins" %in% installed.packages()) {
  remotes::install_github("allisonhorst/palmerpenguins")
}
library(palmerpenguins)
library(ggplot2)

head(penguins)

ggplot(penguins, aes(x = body_mass_g, y = bill_length_mm, color = species)) + 
  geom_point()
```

### Part 2: SAS

```{r, engine="sashtml", engine.path=sasexe, engine.opts=sasopts, collectcode = T, error = T}
PROC PRINT data = sashelp.fish(obs=5);
RUN;

PROC SGPLOT data=sashelp.fish;
SCATTER x = Weight y = Height / Group = Species;
RUN;
```


### Useful References

- [Customizing ggplot2 graphics](http://r-statistics.co/Complete-Ggplot2-Tutorial-Part2-Customizing-Theme-With-R-Code.html)

- [A group-created slideshow of awful graphics](https://docs.google.com/presentation/d/1ixRsGI8BWgT2zA5AbB61eMLUbJgObvBCZt-OnZDzvX4/edit?usp=sharing)

- [PROC SGPLOT procedure options](https://documentation.sas.com/?docsetId=grstatproc&docsetTarget=p1lcbd3lhs3t3bn1jk6d8sjt2yqx.htm&docsetVersion=9.4&locale=en)


